Sampling Strategies in Decisions from Experience

### 2022-01-19 manuscript.Rmd
Renaming of sections and change of figure options

@@

header-includes:
   - \usepackage[capposition=top]{floatrow}
   - \renewcommand\author[1]{}
   - \renewcommand\affiliation[1]{}
   - \authorsnames[1]{Linus Hof}
   - \authorsaffiliations{{Department of Psychology, Heidelberg University}}
---

```{r}
# load packages
pacman::p_load(tidyverse, 
               knitr, 
               ggbeeswarm,
               scico, # for scientific color palettes
               papaja)
```

@@

Often, however, the properties used by the mind to make a choice are not readily described and cannot be directly evaluated [cf. @brunswikRepresentativeDesignProbabilistic1955; see @brunswikPerceptionRepresentativeDesign1956, for a detailed treatment]; rather, proper descriptions must be inferred from whatever information the inputs provide about the properties of choice alternatives, using transformation rules, statistical generalizations, or both.
Since how the inputs are represented determines the information that is made explicit and the costs at which certain operations can be carried out on that information, the feasibility and ease of these inferences depends greatly on the choice of *representation* [@marrVisionComputationalInvestigation1982; see also, e.g., @gigerenzerHowImproveBayesian1995; @griffithsProbabilisticModelsCognition2010; @kempStructuredStatisticalModels2009].
Hence, to explain how the mind makes a selection among choice alternatives, one needs to understand the output controlling properties, i.e., the applied selection criteria, the representation that is used to describe these properties, and the costs of inferring the descriptions in that representation, including their susceptibility to distortions from uncertainty or noise. The process is schematically shown in Figure 1.  
\

@@

# Section 1: Selection Criteria / Properties of Choice Alternatives / Sampling

By defining decision making as an information processing problem whose solution requires a selection among choice alternatives of the form described in (1), much weight is given to the properties of choice alternatives---whatever those may be for the moment---since it is only through them that any necessary distinction between the alternatives or collections thereof can be made. 
To stress this point, consider the power set $\mathcal{P}(I)$, which is the set of all possible subsets of $I$ (see example in Figure 2).
\

@@

Each of the subsets $I' \in \mathcal{P}(I)$ corresponds to a possible distinction between the choice alternatives in $I$ that could result from assessing for each alternative whether it possesses an arbitrary property (collection) or not [cf. @jostMathematicalConcepts2015, p. 16]. 
The concept of the power set thereby illustrates how different selection criteria can alter the consequential choice.
That is, in rather abstract terms, selection criteria are statements regarding the properties of choice alternatives that the mind evaluates as either true or false during the decision making process; accordingly, the subset $I'$ only contains the alternatives for which the statement is evaluated to be true.
It follows that if there is some variance in the properties across choice alternatives and the properties are not perfectly correlated, changing the selection criteria can eventually change the subset $I'$.
<!-- Of course, the whole argument rests on the assumption that ... -->

## What Are the Properties of Choice Alternatives and the Selection Criteria?

The argument so far has been quite general and abstract, since there is no complete answer to the question raised above this paragraph---there is, of course, an uncountable number of decisions and so are there choice alternatives, properties, and selection criteria.
Therefore, in the present paper, *prospects* are used as a proxy for the variety of choice alternatives the mind is required to evaluate in natural world decisions. 
Consequently, the argument continues to proceed at an abstract level: with choices between at least two prospects, the current paper considers a case which is counterfactual, since it is freed from many varieties and complexities that the mind is normally confronted with. 
The case is, however, quite paradigmatic in research on decision theory and there are good reasons for that [cf. @knightRiskUncertaintyProfit1921, especially Chapter 1 for a brief discussion closely related to the current subject]. 
Moreover, the case suffices to answer the above question and to subsequently address the induction and representation problem, at least within the stated limitations.

@@

## Inferring Property Descriptions by Sampling

@@

# Section 2: Underweighting Pattern / Hills and Hertwig (2010) / Induction

# Section 3: Representation

# Section 4: Simulation Study

@@

cpt_long %>% 
  filter(parameter == "gamma" | parameter == "delta") %>% 
  ggplot(aes(x = s, y = mean)) +
  geom_density_2d_filled(contour_var = "ndensity") + 
  scale_fill_scico_d(palette = "buda") +
  scale_color_scico_d(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(limits = c(0,2), breaks = seq(0, 2, .2)) +
  labs(x ="Switching Probability",
       y = "Parameter Value",  
       fill="Density") + 
  facet_grid(parameter~a) + 
  theme_apa()
```

@@

cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x ="Experienced Probability",
       y = "Probability Weighting", 
       color="Switching Probability") + 
  theme_minimal()

@@

cpt_wide %>%
  ggplot(aes(x = s, y = alpha)) +
  geom_density_2d_filled(contour_var = "ndensity") + 
  scale_fill_scico_d(palette = "buda") +
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(limits = c(0,2), breaks = seq(0, 2, .2)) +
  labs(x ="Switching Probability",
       y = "Parameter Value",  
       fill="Density") + 
  facet_wrap(~a, nrow = 1) + 
  theme_minimal()
```

@@

cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) + 
  labs(title = "Value Function",
       x ="Objective Value",
       y = "Subjective Value", 
       color="Switching Probability") + 
  theme_minimal()

@@

# Section 5: Discussion and Conclusion

@@

<div id="refs"></div>

# Appendix

>>>>

affiliation:
  - id            : "1"
    institution   : "Heidelberg University"

@@

header-includes:
   - \usepackage[capposition=top]{floatrow}
---

```{r}
# load packages
pacman::p_load(tidyverse, 
               knitr, 
               ggbeeswarm,
               scico, # for scientific color palettes
               papaja)

knitr::opts_chunk$set(fig.pos = "tb")
```

@@

Often, however, the properties used by the mind to make a choice are not readily described and cannot be directly evaluated [cf. @brunswikRepresentativeDesignProbabilistic1955; see @brunswikPerceptionRepresentativeDesign1956, for a detailed treatment]; rather, proper descriptions must be inferred from whatever information the inputs provide about the properties of choice alternatives, using transformation rules, statistical generalizations, or both.
Since how the inputs are represented determines the information that is made explicit and the costs at which certain operations can be carried out on that information, the feasibility and ease of these inferences depends greatly on the choice of *representation* [@marrVisionComputationalInvestigation1982; see also, e.g., @gigerenzerHowImproveBayesian1995; @griffithsProbabilisticModelsCognition2010; @kempStructuredStatisticalModels2009].
Hence, to explain how the mind makes a selection among choice alternatives, one needs to understand the output controlling properties, i.e., the applied selection criteria, the representation that is used to describe these properties, and the costs of inferring the descriptions in that representation, including their susceptibility to distortions from uncertainty or noise. The process is schematically shown in Figure 1.

@@

# Inputs to the Decision Making Process and Property Descriptions

By defining decision making as an information processing problem whose solution requires a selection among choice alternatives of the form described in (1), much weight is given to the properties of choice alternatives---whatever those may be for the moment---since it is only through them that any necessary distinction between the alternatives or collections thereof can be made. 
To stress this point, consider the power set $\mathcal{P}(I)$, which is the set of all possible subsets of $I$ (see example in Figure 2). 
Each of the subsets $I' \in \mathcal{P}(I)$ corresponds to a possible distinction between the choice alternatives in $I$ that could result from assessing for each alternative whether it possesses an arbitrary property (collection) or not [cf. @jostMathematicalConcepts2015, p. 16].

@@

The concept of the power set thereby illustrates how different selection criteria can alter the consequential choice.
That is, in rather abstract terms, selection criteria are statements regarding the properties of choice alternatives that the mind evaluates as either true or false during the decision making process; accordingly, the subset $I'$ only contains the alternatives for which the statement is evaluated to be true.
It follows that if there is some variance in the properties across choice alternatives and the properties are not perfectly correlated, changing the selection criteria can eventually change the subset $I'$.

## What Are the Choice Alternatives and Their Properties?

The argument so far has been quite general and abstract, since there is no complete answer to the question raised above this paragraph---there is, of course, an uncountable number of decisions and so are there choice alternatives and properties.
Therefore, in the present paper, *prospects* are used as a proxy for the variety of choice alternatives the mind is required to evaluate in natural world decisions. 
Consequently, the argument continues to proceed at an abstract level: with choices between at least two prospects, the current paper considers a case which is counterfactual, since it is freed from many varieties and complexities that the mind is normally confronted with. 
The case is, however, quite paradigmatic in research on decision theory and there are good reasons for that [cf. @knightRiskUncertaintyProfit1921, especially Chapter 1 for a brief discussion closely related to the current subject]. 
Moreover, the case suffices to answer the above question and to subsequently address the induction and representation problem, at least within the stated limitations.

@@

## Inferring Property Descriptions from Sampling Data

@@

# Evaluation of Choice Alternatives and Outputs of the Decision Making Process

# Sampling Strategies as Representation Systems

# Simulation Study

@@

cpt_long %>% 
  filter(parameter == "gamma" | parameter == "delta") %>% 
  ggplot(aes(x = s, y = mean)) +
  geom_density_2d_filled(contour_var = "ndensity") + 
  scale_fill_scico_d(palette = "grayC") +
  scale_color_scico_d(palette = "grayC") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(limits = c(0,2), breaks = seq(0, 2, .2)) +
  labs(x ="Switching Probability",
       y = "Parameter Value",  
       fill="Density") + 
  facet_grid(parameter~a) + 
  theme_apa()
```

@@

cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "grayC") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x ="Experienced Probability",
       y = "Probability Weighting", 
       color="Switching Probability") + 
  theme_minimal()

@@

cpt_wide %>%
  ggplot(aes(x = s, y = alpha)) +
  geom_density_2d_filled(contour_var = "ndensity") + 
  scale_fill_scico_d(palette = "grayC") +
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(limits = c(0,2), breaks = seq(0, 2, .2)) +
  labs(x ="Switching Probability",
       y = "Parameter Value",  
       fill="Density") + 
  facet_wrap(~a, nrow = 1) + 
  theme_minimal()
```

@@

cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(aes(color = s)) +
  scale_color_scico(palette = "grayC") + 
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) + 
  labs(title = "Value Function",
       x ="Objective Value",
       y = "Subjective Value", 
       color="Switching Probability") + 
  theme_minimal()

@@

# Discussion and Conclusion

@@
